{\datastorr}: A package for distributing dynamic and versioned data

### 2016-05-31 ms.tex
changed intro paragraph

\newcommand{\livingdata}{\texttt{living\_data}}
\newcommand{\smurl}[1]{{\footnotesize\url{#1}}}
\usepackage{graphicx}

\title{{\livingdata}: A package for distributing dynamic and versioned
data}
\author{Richard G. FitzJohn$^1$, and fans$^2$}
\affiliation{
$^1$ Somewhere in the UK\\
$^2$ Everywhere else\\

@@

\mstype{Applications Note}
\runninghead{The {\livingdata} package}
\keywords{}

@@

In recent years the idea of the data publication has become established
across many different scientific disciplines. Researchers can publish
the data associated with a publication (or even not associated with a
publication) in one of a number of open access locations, such as the
Dryad Data Repository, Figshare, or (less ideally) as supplementary
files to the original publication. These stable data then are freely and
easily available. Researchers get credit for the production (or
assembly) of that data, and the model facilitates future, creative uses
by the research community. As this new model has taken hold, one new
problem that has emerged: many of the most useful datasets are often not
as static as we once throught they were. The static model works very
well for reproducing a \emph{particular} analysis in a \emph{particular}
paper, which is crucial and important for reproducibility.
{[}\textbf{MWP}: I think we should highlight the fact that there is a
range of types of data from one-off studies of a single population at a
single time (where Dryad is a perfectly suitable option) to large
databases (think GenBank) that require expert database management, good
APIs, etc. I think this approach fills a niche somewhere in the middle.
I realize this is mentioned below but think this would perhaps be useful
upfront to set the scope of the problem we want to address.{]}

From a scientific discovery perspective, however, this is not the most
useful model. Many, although not all, of the most useful datasets are
snapshots of existing quantitative knowledge at a given time. For
example, meta-analyses are built on the existing data for a given
question at a given time. When a new study is published the static
dataset from the last meta-analysis immediately becomes obsolete--to
keep up-to-date that new study should be added. However, if we have a
static data distribution system, this is impossible. Another problem
with the static model is the discovery of small errors. For any much
medium-sized database that is repeatedly used, errors will be found, and
there needs to be some organized way to correct them.

>>>>

\newcommand{\datastorr}{\texttt{datastorr}}
\newcommand{\smurl}[1]{{\footnotesize\url{#1}}}
\usepackage{graphicx}

\title{{\datastorr}: A package for distributing dynamic and versioned
data}
\author{Richard G. FitzJohn$^1$, Daniel S. Falster$^2$, Matthew
  W. Pennell$^3$, and William K. Cornwells$^4$}
\affiliation{
$^1$ Somewhere in the UK\\
$^2$ Everywhere else\\
$^3$ Department of Zoology and Biodiversity Research Centre,
University of British Columbia, Vancouver, B.C. V6T 1Z4 Canada\\

@@

\mstype{Applications Note}
\runninghead{The {\datastorr} package}
\keywords{}

@@

Data is quickly becoming a first-class scientific product. Funding bodies, publishers, and scientific social norms are increasingly both requiring data to be published alongside research publications and recognizing the value of producing and sharing data for its own sake. This has been a rapid, and exciting, sea change and we as a community are still catching up and trying to discern how best to disseminate data and credit data creators and aggregators. One thing that is clear is that there is not a one-size-fits-all solution to this problem as data is incredibly heterogeneous. On one end of the spectrum are massive, centralized repositories that are constantly being added to; this is a great solution to storing and distributing relatively standardized data such as sequences (e.g., \texttt{GenBank}, \texttt{SRA}). On the other end are repositories for storing a wide variety of static, stand-alone datasets (\texttt{Dryad}, \texttt{Figshare}, and less optimally, as supplements to research papers. The static model works very well for reproducing a \emph{particular} analysis in a \emph{particular} paper, which is crucial and important for reproducibility. While both of these classes of data are well-served by current infrastructure, we argue that there is a middle ground that does not fit either of these paradigms very well.

Consider the case of the data from an ecological meta-analysis. It is only a snapshot of our existing knowledge at a given time. When a new study is published the static dataset from the last meta-analysis immediately becomes obsolete --- to keep up-to-date that new study should be added. However, if we have a static data distribution system, this is impossible. Another problem with the static model is the discovery of small errors. For any much medium-sized database that is repeatedly used, errors will be found, and there needs to be some organized way to correct them. And while a centralized GenBank-esque repository would be useful for updating or adding to the data from this meta-analysis, the infrastructure investment (designing maintaining, and hosting the database) is clearly overkill for most projects and is a very substantial barrier to entry.
